---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dbw: Doubly Robust Distribution Balancing Weighting Estimation

<!-- badges: start -->
<!-- badges: end -->

**dbw** implements the doubly robust distribution balancing weighting proposed by Katsumata (2024) <doi:10.1017/psrm.2024.23>, which improves the augmented inverse probability weighting (AIPW) by estimating propensity scores with estimating equations suitable for the pre-specified parameter of interest (e.g., the average treatment effects or the average treatment effects on the treated) and estimating outcome models with the estimated inverse probability weights. 

It also implements the covariate balancing propensity score proposed by Imai and Ratkovic (2014) <doi:10.1111/rssb.12027> and the entropy balancing weighting proposed by Hainmueller (2012) <doi:10.1093/pan/mpr025>, which use covariate balancing conditions in propensity score estimation. 

The point estimate of the parameter of interest and its uncertainty as well as coefficients for propensity score estimation and outcome regression are produced using the M-estimation. The same functions can be used to estimate average outcomes in missing outcome cases.


## How to Cite
<font size="4"> 
Katsumata, Hiroto. 2024. "How Should We Estimate Inverse Probability Weights with Possibly Misspecified Propensity Score Models?" *Political Science Research and Methods*.
</font>

## Installation
<!-- 
You can install the release version of **dbw** from [CRAN](https://CRAN.R-project.org/dbw):
``` r
install.packages("dbw")
```
-->

You can install the development version of **dbw** from [GitHub](https://github.com/hirotokatsumata/dbw):
``` r
devtools::install_github("hirotokatsumata/dbw")
```

## Example
Simulation from Kang and Shafer (2007) and Imai and Ratkovic (2014)

### Make a toy data set
```{r toydata}
library(dbw)
# ATE estimation
# Make a toy data set
# True ATE is 10
tau <- 10
set.seed(12345)
n <- 1000
X <- matrix(stats::rnorm(n * 4, mean = 0, sd = 1), nrow = n, ncol = 4)
prop <- 1 / (1 + exp(X[, 1] - 0.5 * X[, 2] + 0.25 * X[, 3] + 0.1 * X[, 4]))
treat <- rbinom(n, 1, prop)
y <- 210 + 27.4 * X[, 1] + 13.7 * X[, 2] + 13.7 * X[, 3] + 13.7 * X[, 4] + 
     tau * treat + stats::rnorm(n = n, mean = 0, sd = 1)
ybinom <- (y > 210) + 0
df0 <- data.frame(X, treat, y, ybinom)
colnames(df0) <- c("x1", "x2", "x3", "x4", "treat", "y", "ybinom")

# Variables for a misspecified model
Xmis <- data.frame(x1mis = exp(X[, 1] / 2), 
                   x2mis = X[, 2] * (1 + exp(X[, 1]))^(-1) + 10,
                   x3mis = (X[, 1] * X[, 3] / 25 + 0.6)^3, 
                   x4mis = (X[, 2] + X[, 4] + 20)^2)

# Data frame and formulas for propensity score estimation
df <- data.frame(df0, Xmis)
formula_ps_c <- stats::as.formula(treat ~ x1 + x2 + x3 + x4)
formula_ps_m <- stats::as.formula(treat ~ x1mis + x2mis + x3mis + x4mis)

# Formula for a misspecified outcome model
formula_y <- stats::as.formula(y ~ x1mis + x2mis + x3mis + x4mis)
```

### Correct propensity score model

```{r correct model}
# Distribution balancing weighting with normalization and without regularization
fitdbwc <- dbw(formula_y = formula_y, formula_ps = formula_ps_c, 
               estimand = "ATE", method = "dbw",
               method_y = "wls", data = df, normalize = TRUE, 
               vcov = TRUE, lambda = 0, weights = NULL, clevel = 0.95)
fitdbwc
summary(fitdbwc)

# Covariate balancing weighting function without regularization
fitcbwc <- dbw(formula_y = formula_y, formula_ps = formula_ps_c, 
               estimand = "ATE", method = "cb",
               method_y = "wls", data = df, normalize = TRUE, 
               vcov = TRUE, lambda = 0, weights = NULL, clevel = 0.95)
summary(fitcbwc)

# Entropy balancing weighting function without regularization
fitebwc <- dbw(formula_y = formula_y, formula_ps = formula_ps_c, 
               estimand = "ATE", method = "eb",
               method_y = "wls", data = df, normalize = TRUE, 
               vcov = TRUE, lambda = 0, weights = NULL, clevel = 0.95)
summary(fitebwc)

# Standard logistic regression
fitmlec <- dbw(formula_y = formula_y, formula_ps = formula_ps_c, 
               estimand = "ATE", method = "mle",
               method_y = "wls", data = df, normalize = FALSE, 
               vcov = TRUE, lambda = 0, weights = NULL, clevel = 0.95)
summary(fitmlec)

# Distribution balancing weighting without normalization and without regularization
fitdbwcnn <- dbw(formula_y = formula_y, formula_ps = formula_ps_c, 
                 estimand = "ATE", method = "dbw",
                 method_y = "wls", data = df, normalize = FALSE, 
                 vcov = TRUE, lambda = 0, weights = NULL, clevel = 0.95)
summary(fitdbwcnn)
```

### Misspecified propensity score model
```{r misspecified model}
# Distribution balancing weighting with normalization and without regularization
fitdbwm <- dbw(formula_y = formula_y, formula_ps = formula_ps_m, 
               estimand = "ATE", method = "dbw",
               method_y = "wls", data = df, normalize = TRUE, 
               vcov = TRUE, lambda = 0, weights = NULL, clevel = 0.95)
summary(fitdbwm)

# Covariate balancing weighting function without regularization
fitcbwm <- dbw(formula_y = formula_y, formula_ps = formula_ps_m, 
               estimand = "ATE", method = "cb",
               method_y = "wls", data = df, normalize = TRUE, 
               vcov = TRUE, lambda = 0, weights = NULL, clevel = 0.95)
summary(fitcbwm)

# Entropy balancing weighting function without regularization
fitebwm <- dbw(formula_y = formula_y, formula_ps = formula_ps_m, 
               estimand = "ATE", method = "eb",
               method_y = "wls", data = df, normalize = TRUE, 
               vcov = TRUE, lambda = 0, weights = NULL, clevel = 0.95)
summary(fitebwm)

# Standard logistic regression
fitmlem <- dbw(formula_y = formula_y, formula_ps = formula_ps_m, 
               estimand = "ATE", method = "mle",
               method_y = "wls", data = df, normalize = FALSE, 
               vcov = TRUE, lambda = 0, weights = NULL, clevel = 0.95)
summary(fitmlem)

# Distribution balancing weighting without normalization and without regularization
fitdbwmnn <- dbw(formula_y = formula_y, formula_ps = formula_ps_m, 
                 estimand = "ATE", method = "dbw",
                 method_y = "wls", data = df, normalize = FALSE, 
                 vcov = TRUE, lambda = 0, weights = NULL, clevel = 0.95)
summary(fitdbwmnn)

# Distribution balancing weighting with normalization and with regularization
# Standardization
res_std_comp <- std_comp(formula_y = formula_y, 
                         formula_ps = formula_ps_m, 
                         estimand = "ATE", method_y = "wls", 
                         data = df, std = TRUE,
                         weights = NULL)
# Estimation
fitdbwmr <- dbw(formula_y = formula_y, 
                formula_ps = res_std_comp$formula_ps, 
                estimand = "ATE", method = "dbw", method_y = "wls",
                data = res_std_comp$data, normalize = TRUE, 
                vcov = TRUE, lambda = 0.01, weights = res_std_comp$weights, 
                clevel = 0.95)
summary(fitdbwmr)

# Covariate balancing weighting function with an estimating equation 
#  for the original covariate balancing propensity score method
fitcbwmcmb <- dbw(formula_y = formula_y, formula_ps = formula_ps_m, 
                  estimand = "ATEcombined", method = "cb",
                  method_y = "wls", data = df, normalize = TRUE, 
                  vcov = TRUE, lambda = 0, weights = NULL, clevel = 0.95)
summary(fitcbwmcmb)


# Formula for a misspecified outcome model for the GAM
library(mgcv)
formula_y_gam <- stats::as.formula(y ~ s(x1mis) + s(x2mis) + 
                                       s(x3mis)+ s(x4mis))

# Distribution balancing weighting with the GAM
fitdbwmg <- dbw(formula_y = formula_y_gam, formula_ps = formula_ps_m, 
                estimand = "ATE", method = "dbw",
                method_y = "gam", data = df, normalize = TRUE, 
                vcov = TRUE, lambda = 0, weights = NULL, clevel = 0.95)
summary(fitdbwmg)
```


### Binary outcome case
```{r binary outcome}
# Empirically correct ATE
ybinom_t <- (y + (1 - treat) * 10 > 210) + 0
ybinom_c <- (y - treat * 10 > 210) + 0
ATEbinom <- mean(ybinom_t - ybinom_c)
ATEbinom

# Formula for a misspecified binary outcome model
formula_y_bin <- stats::as.formula(ybinom ~ x1mis + x2mis + x3mis + x4mis)

# Distribution balancing weighting for the binary outcome
fitdbwmbin <- dbw(formula_y = formula_y_bin, formula_ps = formula_ps_m, 
                  estimand = "ATE", method = "dbw",
                  method_y = "logit", data = df, normalize = TRUE, 
                  vcov = TRUE, lambda = 0, weights = NULL, clevel = 0.95)
summary(fitdbwmbin)
```


### Horvitz-Thompson and Hajek estimators
```{r HT estimator}
# Standard logistic regression with the Horvitz-Thompson estimator
fitmlem_ht <- dbw(formula_y = y ~ 0, formula_ps = formula_ps_m, 
                  estimand = "ATE", method = "mle",
                  method_y = "wls", data = df, normalize = FALSE, 
                  vcov = TRUE, lambda = 0, weights = NULL, clevel = 0.95)
summary(fitmlem_ht)

# Standard logistic regression with the Hajek estimator
fitmlem_hj <- dbw(formula_y = y ~ 1, formula_ps = formula_ps_m, 
                  estimand = "ATE", method = "mle",
                  method_y = "wls", data = df, normalize = FALSE, 
                  vcov = TRUE, lambda = 0, weights = NULL, clevel = 0.95)
summary(fitmlem_hj)
```


## Covariate balance plot
```{r covariate balance}
# Distribution balancing weighting without regularization
oldpar <- par(no.readonly = TRUE) # Just for adjusting plot margins
par(mar = c(5.1, 5.1, 4.1, 2.1)) # Just for adjusting plot margins
plot(fitdbwm, addcov = ~ x1 + x2 + x3 + x4, threshold = c(0.1, 0.2))
par(oldpar)

# Display the covariate balance matrix
cb <- plot(fitdbwm, addcov = ~ x1 + x2 + x3 + x4, plot = FALSE)
cb
```
